Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-07-17 chapters/ch_appendix.tex
backup
Fixed Appendix
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.
Added bills edits working on appendix figures

\subsection{$\Delta$ Resonance Production and Decay Channels}
\label{appen:deltadecay}

@@

\begin{equation}
\begin{split}
& p + p \rightarrow \sqrt{\frac{5}{6}}(\pi^+ + p + n) - \sqrt{\frac{1}{6}}(\pi^0 + p + p) \\
& n + p \rightarrow \sqrt{\frac{1}{6}}(\pi^+ + n + n) + \sqrt{\frac{2}{3}}(\pi^0 + n + p) + \sqrt{\frac{1}{6}}(\pi^- + p + p)  \\
& n + n \rightarrow \sqrt{\frac{1}{6}}(\pi^0 + n + n) - \sqrt{\frac{5}{6}}(\pi^- + n + p) 
\end{split}
\label{eq:deltaProduction}
\end{equation}

@@

\subsection{Runs analyzed in this data}

@@

\subsection{Pion Yield Theory}
\label{tb:pionyieldTheory}

@@

\section{Cut Variation Analysis}
\label{sec:cutvar}
The particular track quality cuts described in Section~\ref{sec:qualitycut} are there to reduced the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra. The best set of values were found for all the data sets which include the charged particle multiplicity of the event, dOCA, and the number of clusters cut. By varying the cuts in both the data and the MC efficiency, we can evaluate the analysis to see if there are any systematic dependencies on the observable we plot. But varying the cuts undoutably means as the cut get tighter, and less data is included, the statistical error bar will increase making it difficult to compare to more loose cuts. This is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}. Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts and the uncorrelated error bars, then there is some systematic trend that either exists due to physics or some error or miscalculation in the analysis. It is usually not recommended to estimated systematic error bars using this method, but sometimes is the only way.

Here we will discuss the particular analysis using the total pion ratio as the example. Figure~\ref{fig:totalRatioError} shows the total pion as a function of several cut variations, where one cut is varied and all the others are held constant. Three cuts are of particular interest, the event multiplicity, dOCA, and the number of cluster cut. Figure~\ref{fig:clustTotalRatio} shows the variation as the number of cluster cut is varied. The default cut used is > 20 clusters which is represented by the middle point. The red bars represent the statistical error of this default value. The error bars on the other points are the uncorrelated error bars as described by the prescription in \cite{dataAnalysis}. Here it is clear that there is a systematic bias in the pion ratio as a function

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pimTotal}
         \caption{No efficiency correction.}
         \label{fig:multCutVar}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pimTotal}
         \caption{Efficiency corrected.}
         \label{fig:multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:cutVar}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pipTotal}
         \caption{No efficiency correction.}
         \label{fig:pip_multCutVar}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pipTotal}
         \caption{Efficiency corrected.}
         \label{fig:pip_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:cutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_pipTotal}
         \caption{Before efficiency correction.}
         \label{fig:pip_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_pipTotal}
         \caption{After efficiency correction.}
         \label{fig:pip_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the number of cluster cut of the tracks.}
\label{fig:pim_clustVar}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_TotalRatio}
         \caption{No efficiency correction.}
         \label{fig:ratio_multCutVar}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_TotalRatio}
         \caption{Efficiency corrected.}
         \label{fig:ratio_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:cutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_TotalRatio}
         \caption{Before efficiency correction.}
         \label{fig:ratio_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_TotalRatio}
         \caption{After efficiency correction.}
         \label{fig:ratio_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the number of cluster cut of the tracks.}
\label{fig:pim_clustVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_NoEC_TotalRatio}
         \caption{No efficiency correction.}
         \label{fig:ratio_cutVarPOCA_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_TotalRatio}
         \caption{Efficiency corrected.}
         \label{fig:ratio_cutVarPOCA_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the dOCA cut.}
\label{fig:pip_cutVarPOCA}
\end{figure}

\chapter{Second appendix}

>>>>

%\setcounter{figure}{0} % Perhaps
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\thetable}{A\arabic{figure}}

\section{$\Delta$ Resonance Production and Decay Channels}
\label{appen:deltadecay}

@@

\begin{equation}
\begin{split}
& p + p \rightarrow \sqrt{\frac{5}{6}}(\pi^+ + p + n) - \sqrt{\frac{1}{6}}(\pi^0 + p + p) \\
& n + p \rightarrow \sqrt{\frac{1}{6}}(\pi^+ + n + n) + \sqrt{\frac{2}{3}}(\pi^0 + n + p) + \sqrt{\frac{1}{6}}(\pi^- + p + p)  \\
& n + n \rightarrow \sqrt{\frac{1}{6}}(\pi^0 + n + n) - \sqrt{\frac{5}{6}}(\pi^- + n + p) 
\end{split}
\label{eq:pionProduction}
\end{equation}

@@

\section{Runs analyzed in this data}

@@

\section{Pion Yield Theory}
\label{tb:pionyieldTheory}

@@

\section{Cut Variation Analysis}
\label{sec:cutvar}

We employ a variety of track quality cuts described in Section~\ref{sec:qualitycut} to reduce the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra. The best set of values were found for all the data sets which include the charged particle multiplicity of the event, the distance of closest approach to the vertex (DOCA), and the minimum number of clusters required for a track to be analyzed. By varying the cuts in both the data and in our MC efficiency calculations, we can evaluate the analysis to see how stable are the results with respect to these cuts. It is intuitive that while varying the cuts we will find that as the cut get tighter, and less data is included, the statistical error bar will increase. This problem is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}. Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts, with respects to the uncorrelated error bars, then there truly is some systematic trend that either exists due to physics or some error or miscalculation in the analysis. It is usually not recommended to estimated systematic error bars using this method, but sometimes it is the only way.

Here we will discuss the particular analysis using the total pion ratio as the example. Figure~\ref{fig:totalRatioError} shows the total pion ratio as a function of several cut variations, where one cut is varied and all the others are held constant. Three cuts are of particular interest, the event multiplicity, DOCA, and the number of cluster cut.

 Figure~\ref{fig:pim_clustVar_NoEC} shows the variation as the number of clusters cut is varied before the efficiency analysis. The default cut for the number of clusters in a track is $N_c$ > 20 clusters which is represented by the middle point. The red bars represent the statistical error of this default value. The error bars on the other points are the uncorrelated error bars as described by the prescription in \cite{dataAnalysis}. Here it is clear that there is a systematic bias in the $\pi^-$ yield as we increase the number of cluster cut. This expected since cutting on the number of clusters eventually throws away short tracks that may be in the large angular regions of acceptance, thus biasing the yield. As seen in Figure~\ref{fig:pim_clustVar_EC}, this is corrected mostly be the efficiency analysis, though there seems to be some residual over correction for above 20 clusters and under correction under it, but still approximately within the statistical error bar of the default cut value of 20.

When looking at the variation of the DOCA cut, we see an even better story. Before the efficiency correction in Figure~\ref{fig:pim_cutVarPOCA_NoEC}, there is very little systematic dependence, except maybe for higher values of DOCA. After efficiency correction we really see no significant variation and can safely claim there are no systematic dependencies on this particular cut.

In Figure~\ref{fig:pim_multCutVar}, we see  a significant systematic variation of the $\pi^-$ yield as a function of the event track multiplicity. Since the event multiplicity is related to how central of a collision we had, this is purely a physics effect that is expected. This is because as the collision becomes more central, more nucleons participate in the collision due to geometric overlap. As the participating nucleons increases, there is a higher chance to make more pions, therefore the yield scales with the track multiplicity, which is ultimately related to the impact parameter as discussed in Section~\ref{sec:impact}. Therefore we should not associate any systematic error due to this effect.

There is a similar discussion that is made for the $\pi^+$ yield in the same cut variations; for the multiplicity cut in Figure~\ref{fig:pip_multCutVar}, the DOCA cut in Figure~\ref{fig:pip_cutVarPOCA}, and the cluster cut in Figure~\ref{fig:pip_clustVar}.

The analysis becomes more interesting when analyzing the total pion ratio for the same cut variation analysis performed above. Where the total pion ratio $\pi^-/\pi^+$ is plotted as a function of the multiplicity cut in Figure~\ref{fig:ratio_multCutVar}, the DOCA cut in Figure~\ref{fig:ratio_cutVarPOCA}, and the cluster cut in Figure~\ref{fig:ratio_clustVar}. Neglecting the multiplicity cut variation, which we know is derived from physical effects, it appears the pion ratio cancels out most of the systematic variations we saw previously in the single particle yields. This was expected, as the ratio tends to cancel systematic effects in the detector system. This is definitive proof of this in action. The systematic trends are less significant than the statistical error bars given in the red region for the default cuts, except maybe for a cluster cut of $N_c > 26$ which is already at the extreme end of the cut, since we know that it cuts into the particle acceptance.

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pimTotal}
         \caption{No efficiency correction.}
         \label{fig:pim_multCutVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pimTotal}
         \caption{Efficiency corrected.}
         \label{fig:pim_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:pim_multCutVar}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pipTotal}
         \caption{No efficiency correction.}
         \label{fig:pip_multCutVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pipTotal}
         \caption{Efficiency corrected.}
         \label{fig:pip_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:pip_multCutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_pipTotal}
         \caption{Before efficiency correction.}
         \label{fig:pip_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_pipTotal}
         \caption{After efficiency correction.}
         \label{fig:pip_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$) when varying the number of cluster cut of the tracks.}
\label{fig:pip_clustVar}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_TotalRatio}
         \caption{No efficiency correction.}
         \label{fig:ratio_multCutVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{132Sn_Multiplicity_TotalRatio}
         \caption{Efficiency corrected.}
         \label{fig:ratio_multCutVar}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
\label{fig:ratio_multCutVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_NoEC_TotalRatio}
         \caption{Before efficiency correction.}
         \label{fig:ratio_clustVar_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Clusters_TotalRatio}
         \caption{After efficiency correction.}
         \label{fig:ratio_clustVar_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the number of cluster cut of the tracks.}
\label{fig:ratio_clustVar}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_NoEC_TotalRatio}
         \caption{No efficiency correction.}
         \label{fig:ratio_cutVarPOCA_NoEC}
     \end{subfigure}
     \hfill
    \begin{subfigure}[b]{\textwidth}
         \centering
         \includegraphics[width=\textwidth]{POCA_TotalRatio}
         \caption{Efficiency corrected.}
         \label{fig:ratio_cutVarPOCA_EC}
     \end{subfigure}
     \hfill
\caption{Y($\pi^+$)/Y($\pi^-$) when varying the dOCA cut.}
\label{fig:ratio_cutVarPOCA}
\end{figure}

\begin{figure}[!htb]
	\centering
    \includegraphics[width=\textwidth]{clusterSpectraSn132}
	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut.}
	\label{fig:clusterSpectralSn132}
\end{figure}

\begin{figure}[!htb]
	\centering
    \includegraphics[width=\textwidth]{pocaSpectraSn132}
	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut.}
	\label{fig:docaSpectralSn132}
\end{figure}

\begin{figure}[!htb]
	\centering
    \includegraphics[width=\textwidth]{multSpectraSn132}
	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}Sn$ cut.}
	\label{fig:multSpectralSn132}
\end{figure}

\begin{figure}[!htb]
	\centering
    \includegraphics[width=\textwidth]{clusterSpectraSn108}
	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut.}
	\label{fig:clusterSpectralSn132}
\end{figure}

\begin{figure}[!htb]
	\centering
    \includegraphics[width=\textwidth]{pocaSpectraSn108}
	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut.}
	\label{fig:docaSpectralSn132}
\end{figure}

\begin{figure}[!htb]
	\centering
    \includegraphics[width=\textwidth]{multSpectraSn108}
	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{108}Sn$ cut.}
	\label{fig:multSpectralSn132}
\end{figure}

%\chapter{Second appendix}

### 2020-07-17 chapters/ch_experiment.tex
Added bills edits working on appendix figures

\chapter{Experiment}

@@

%explain in breif the basics of a time projection chamber 
A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector. TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}. In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this thesis. Figure~\ref{fig:tpcPrinciple} depicts the interior of a gaseous TPC and its field cage. The field cage holds the detector gas and also defines a constant electric field that is typically aligned with the magnetic field of a magnet that surrounds the TPC. In this Spirit TPC, both these electric and magnetic fields are vertical.

@@

where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction  of a given roaw the PRF can be written as,

@@

\begin{comment}
\subsection{Considerations when constructing a TPC}
Several considerations went into the construction of the S$\pi$RI TPC which I wish to summarize and document here. All materials and glues of the TPC were selected as low out-gassing materials. Several materials (that are common place in nuclear labs), such as vacuum grease, viton o-rings, all out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time. The organic molecules responsible are difficult to identify exactly, but lists of good and bad materials are well known in the literature from experiments. If a material we wished to used was not on these lists we placed the material in a clean chamber with the counter gas and flowed this counter gas through a small proportional counter making sure the gain did not drop at high collection rates when exposed to a high rate alpha Americium source.

@@

There were two beam drift chambers (BDC) located along the beam pipe upstream of the magnet and after the last focusing quadropole magnet. These beam drift chambers contained two sets of parallel plate avalanche counters (PPAC) which could get the (x,y) coordinates of the beam. Since they were places approximately \SI{1}{\metre} apart, we were able to track the direction of the beam in the beam line using a linear extrapolation. The resolution of the initial angle the beam enters the SAMURAI dipole was estimated to be \SI{0.64}{\milli\radian} and \SI{0.24}{\milli\radian} for the two angles defining the beam vector \cite{jon}. The initial angle, energy, and charge of the beam is then propogated through the magnetic field map until the angle on target is found. The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Section~\ref{sec:beamangle} and Section~\ref{sec:pionSpectra}.

@@

There were several special trigger considerations incorporated into the trigger for the TPC. We required that the gating grid be opened quickly to not lose electrons from portions of the tracks that pass close to the gating grid. If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicty >4 signal, and neither a Katana Veto signal nor an active target signal. This was referred to as the Fast Trigger, and it triggered the opening of the gating grid, which took about 350 ns. Following the initial trigger, we had the possibility of a fast clear signal that could be initiated by a later beam particle traversing the TPC. If the last beam particle came within xx us after the fast trigger, electrons from this beam particle could pass through the gating grid. Such data would be useless, so we closed the gating grid and stopped the data equisition with a fast clear signal. Additional fast clears could also initiated if the Katana trigger box did not generate a trigger. Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.

@@

Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal. The KATANA trigger, Kyoto Muliplicity, and Start Counter together triggered the DAQ.

>>>>

\chapter{Experiment}
A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector. TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}. In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this Thesis, along with the rest of the experimental setup.

@@

%explain in breif the basics of a time projection chamber 
Figure~\ref{fig:tpcPrinciple} depicts the interior of a gaseous TPC and its field cage. The field cage holds the detector gas and also defines a constant electric field that is typically aligned with the magnetic field of a magnet that surrounds the TPC. In this Spirit TPC, both these electric and magnetic fields are vertical.

@@

where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction  of a given row the PRF can be written as,

@@

\begin{comment}
\subsection{Considerations when constructing a TPC}
Several considerations went into the construction of the S$\pi$RI TPC which I wish to summarize and document here. All materials and glues of the TPC were selected as low out-gassing materials. Several materials (that are common place in nuclear labs), such as vacuum grease, Viton o-rings, all out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time. The organic molecules responsible are difficult to identify exactly, but lists of good and bad materials are well known in the literature from experiments. If a material we wished to used was not on these lists we placed the material in a clean chamber with the counter gas and flowed this counter gas through a small proportional counter making sure the gain did not drop at high collection rates when exposed to a high rate alpha Americium source.

@@

There were two beam drift chambers (BDC) located along the beam pipe upstream of the magnet and after the last focusing quadropole magnet. These beam drift chambers contained two sets of parallel plate avalanche counters (PPAC) which could get the (x,y) coordinates of the beam. Since they were places approximately \SI{1}{\metre} apart, we were able to track the direction of the beam in the beam line using a linear extrapolation. The resolution of the initial angle the beam enters the SAMURAI dipole was estimated to be \SI{0.64}{\milli\radian} and \SI{0.24}{\milli\radian} for the two angles defining the beam vector \cite{jon}. The initial angle, energy, and charge of the beam is then propagated through the magnetic field map until the angle on target is found. The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Section~\ref{sec:beamangle} and Section~\ref{sec:pionSpectra}.

@@

There were several special trigger considerations incorporated into the trigger for the TPC. We required that the gating grid be opened quickly to not lose electrons from portions of the tracks that pass close to the gating grid. If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicity >4 signal, and neither a Katana Veto signal nor an active target signal. This was referred to as the Fast Trigger, and it triggered the opening of the gating grid, which took about 350 ns. Following the initial trigger, we had the possibility of a fast clear signal that could be initiated by a later beam particle traversing the TPC. If the last beam particle came within xx us after the fast trigger, electrons from this beam particle could pass through the gating grid. Such data would be useless, so we closed the gating grid and stopped the data acquisition with a fast clear signal. Additional fast clears could also initiated if the Katana trigger box did not generate a trigger. Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.

@@

Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal. The KATANA trigger, Kyoto Multiplicity, and Start Counter together triggered the DAQ.
